---
title: "Getting and Cleaning Data Course Project"

---

##Instructions for project

The purpose of this project is to demonstrate the ability to collect, work with, and clean a data set. The goal is to prepare tidy data that can be used for later analysis.The reqirement is 
1)To submit a tidy data set as described below.
2)To submit a link to a Github repository with your script for performing the analysis. 
3)To submit a code book that describes the variables, the data, and any transformations or work that is performed to clean up the data called CodeBook.md. Also includes a README.md in the repo with the scripts. This repo explains how all of the scripts work and how they are connected.

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:

(http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones)

Here are the data for the project:

(https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip)

**Creating  R script called run_analysis.R that does the following**

Merges the training and the test sets to create one data set.
Extracts only the measurements on the mean and standard deviation for each measurement.
Uses descriptive activity names to name the activities in the data set
Appropriately labels the data set with descriptive variable names.
From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

##Get the data

**1.Download the file and put the file in the data folder**

```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip")
dateDownloaded<-date()
```

**2.Unzip the file**

```{r}
unzip(zipfile="./data/Dataset.zip",exdir="./data")
```
**3.Unzipped files are in the folder UCI HAR Dataset. Get the list of the files**

```{r}
path_rf <- file.path("./data" , "UCI HAR Dataset")
files<-list.files(path_rf, recursive=TRUE)
files
```
See the README.txt file for the detailed information on the dataset. For the purposes of this project, the files in the Inertial Signals folders are not used. The files that will be used to load data are listed as follows:

**test/subject_test.txt**  
**test/X_test.txt**   
**test/y_test.txt**    
**train/subject_train.txt**   
**train/X_train.txt**   
**train/y_train.txt**    

**Read data from the targeted files**


We use Activity, Subject and Features as part of descriptive variable names for data in data frame.

**2.Read data from the files into the variables**

**Read the Activity files**

```{r}
dataActivityTest  <- read.table(file.path(path_rf, "test" , "Y_test.txt" ),header = FALSE)
dataActivityTrain <- read.table(file.path(path_rf, "train", "Y_train.txt"),header = FALSE)
```
**Read the Subject files**
```{r}
dataSubjectTrain <- read.table(file.path(path_rf, "train", "subject_train.txt"),header = FALSE)
dataSubjectTest  <- read.table(file.path(path_rf, "test" , "subject_test.txt"),header = FALSE)
```
**Read Feature files**
```{r}
dataFeaturesTest  <- read.table(file.path(path_rf, "test" , "X_test.txt" ),header = FALSE)
dataFeaturesTrain <- read.table(file.path(path_rf, "train", "X_train.txt"),header = FALSE)
```
The properties of the above varibles
```{r}
str(dataActivityTest)
str(dataActivityTrain)
str(dataSubjectTrain)
str(dataSubjectTest)
str(dataFeaturesTest)
str(dataFeaturesTrain)
```
##Merge the training and the test sets to create one data set

**1.Concatenate the data tables by rows**
```{r}
dataSubject <- rbind(dataSubjectTrain, dataSubjectTest)
dataActivity<- rbind(dataActivityTrain, dataActivityTest)
dataFeatures<- rbind(dataFeaturesTrain, dataFeaturesTest)
```
**2.Set names to variables**
```{r}
names(dataSubject)<-c("subject")
names(dataActivity)<- c("activity")
dataFeaturesNames <- read.table(file.path(path_rf, "features.txt"),head=FALSE)
names(dataFeatures)<- dataFeaturesNames$V2
```

**3.Merge columns to get the data frame Data for all data**
```{r}
dataCombine <- cbind(dataSubject, dataActivity)
Data <- cbind(dataFeatures, dataCombine)
```
**Extracts only the measurements on the mean and standard deviation for each measurement**

**1.Subset Name of Features by measurements on the mean and standard deviation
i.e taken Names of Features with "mean()" or "std()"**

```{r}
subdataFeaturesNames<-dataFeaturesNames$V2[grep("mean\\(\\)|std\\(\\)", dataFeaturesNames$V2)]
```
**2.Subset the data frame Data by seleted names of Features**
```{r}
selectedNames<-c(as.character(subdataFeaturesNames), "subject", "activity" )
Data<-subset(Data,select=selectedNames)
```
**3.Check the structures of the data frame Data**
```{r}
str(Data)
```

##Uses descriptive activity names to name the activities in the data set 

**1.Read descriptive activity names from "activity_labels.txt"**
```{r}

activityLabels <- read.table(file.path(path_rf, "activity_labels.txt"),header = FALSE)
```
**2.Factorize Variable activity in the data frame Data using descriptive activity names**

**3.Check**
```{r}
head(Data$activity,30)
```
Appropriately labels the data set with descriptive variable names
In the former part, variables activity and subject and names of the activities have been labelled using descriptive names.In this part, Names of Feteatures will labelled using descriptive variable names.

*prefix t is replaced by time*
*Acc is replaced by Accelerometer*
*Gyro is replaced by Gyroscope*
*prefix f is replaced by frequency*
*Mag is replaced by Magnitude*
*BodyBody is replaced by Body*
```{r}
names(Data)<-gsub("^t", "time", names(Data))
names(Data)<-gsub("^f", "frequency", names(Data))
names(Data)<-gsub("Acc", "Accelerometer", names(Data))
names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
names(Data)<-gsub("Mag", "Magnitude", names(Data))
names(Data)<-gsub("BodyBody", "Body", names(Data))
```
check
```{r}
names(Data)
```

##Creates a second,independent tidy data set and displays it

In this part,a second, independent tidy data set will be created with the average of each variable for each activity and each subject based on the data set in step 4.

```{r}
library(plyr)
Data2<-aggregate(. ~subject + activity, Data, mean)
Data2<-Data2[order(Data2$subject,Data2$activity),]
write.table(Data2, file = "tidydata.txt",row.name=FALSE)
knit2html("codebook.Rmd")
```

